---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# valetr

`valetr` provides an interface to the [Bank of Canada’s
Valet API](https://www.bankofcanada.ca/valet/docs) —read the API's [terms and
condtions](https://www.bankofcanada.ca/terms/). Some of the data
available include key monetary policy variables —i.e. Total CPI and M
aggregates.

Valet also provides a route to Foreign Exchange Rates in RSS. This route
returns the **most recent** foreign exchange rate —one observation only. This functionality is not included in `valetr`.

## Retrieve key monetary policy variables

The Valet API houses key monetary policy variables series. For instance, it is possible to extract the series for some measures of core inflation:

```{r eval=FALSE}
# getSeriesInfo will return a data.frame with the series info (including link and name)
cpiInfo <- getSeriesInfo(patternGroupLabel = "consumer price index$",
              patternSeriesLabel = "trim|median|common|total cpi$",  
                ignore.case=TRUE)

# now use the series'  link or name to feed getSerieData
# note that getSeriesData accepts multilpe links or names
cpis <- getSeriesData(series=cpiInfo[["series_link"]]) # can also use cpiInfo[["series_name"]]
```

Note that `getSeriesData()` allows for the API's query functionality:

```{r eval=FALSE}
cpis <- getSeriesData(series=cpiInfo[["series_link"]], start_date="2008-07-01", end_date="2012-07-01")
# cpis <- getSeriesData(series=cpiInfo[["series_link"]], recent_years=12L) # last 12 yrs
# cpis <- getSeriesData(series=cpiInfo[["series_link"]], recent_months=12L) # last 12 months
# cpis <- getSeriesData(series=cpiInfo[["series_link"]], recent_weeks=12L) # last 12 weeks
# cpis <- getSeriesData(series=cpiInfo[["series_link"]], recent=12L) # last 12 obs
```

Note that only `start_date` and `end_date` can be used together as query parameters.

## Obtain exchange rate series

You can also retrieve foreign exchange rate series:

```{r eval=FALSE}
fxInfo <- getSeriesInfo(patternGroupLabel = "monthly exchange",
              patternSeriesLabel = "(EUR|GBP|USD)\\/CAD",  
                ignore.case=TRUE)

# now use the series'  link or name to feed getSerieData
# note that getSeriesData accepts multilpe links or names
fx <- getSeriesData(series=fxInfo[["series_link"]]) # can also use fx[["series_name"]]
```

## Installation

`valetr` has yet to be submitted to CRAN.

To install the development version:

``` r
devtools::install_github("jdemello/valetr")
```

## Functions


```
regexSeriesLabel  Matches a `pattern` to the label of one or more series.
                  Returns a `data.frame` with label, link and name

getSeriesInfo     Allows to pattern-match both group and series labels. 
                  This function returns a `data.frame` with a more 
                  complete description of a series. It includes their 
                  group information (name, label, desc.) as well as 
                  the series descriptive components

getSeriesData     Extracts one or more series by specifying their name 
                  or link --vector can contain a mix of both. It also 
                  allows the insertion of the API's query parameters.
```


Read more about the API [here](https://www.bankofcanada.ca/valet/docs#valet_api).

